Software Product Line Analysis Techniques

### 2018-01-03 doc/quals/quals.tex
Liebig's FSE'13 paper

\section{Scalable Analysis of Variable Software}
\cite{Liebig:2013}

@@

>>>>

@@

\section{Scalable Analysis of Variable Software}
In \cite{Liebig:2013}, variability-aware (family-based) static analyses are compared to some sample-based techniques. Two variability-aware C language analyses are implemented and evaluated: type checking and liveness analysis. Three relatively big open source systems were used for evaluation: the Linux kernel, the Busybox codebase, and the OpenSSL library.

In addition to focusing on static analysis as opposed to model checking, there are two main differences between this work and that of \cite{Apel:2013}. First, the variability-aware analyses are applied to annotative SPLs (using the CPP to incorporate variability). Because annotative SPLs and more permissive than generative SPLs in terms of syntactic variability, some building blocks had to be purposely designed to be variability-aware. Those include variability-aware Abstract Syntax Trees (ASTs), Control Flow Graphs (CFGs) and symbol tables. In addition, a dedicated variability-aware parser (incorporating CPP logic) is used.

The second difference with respect to \cite{Apel:2013} is that the sampling techniques used in the comparison are more elaborate. In addition to pair-wise sampling (covered in \cite{Apel:2013}), single configuration, random and coverage-based sampling were also evaluated.

Similar to \cite{Apel:2013}, the main principles guiding the design of the variability-aware analyses are late splitting and early joining. These principles maximize reuse of intermediate analysis information across different variants, eventually resulting in variability-aware analyses outperformaing brute-force per-product analysis by orders of magnitude. A third principle implicit in \cite{Apel:2013} but explicitly spelled out here is local variability representation. This is the mechanism of maintaining variability within intermediate analysis values.

The evaluation presented in the paper justifies the hypotheses of the authors: the implemented variability-aware analyses under-perform single configuration analyses, but outperform the analysis of pair-wise samples. In addition, in some cases variability-aware analyses also outperform coverage-based analysis. More importantly, variability-aware analyses are complete in terms of covering all valid feature combinations (i.e. all valid products), while all the sampling techniques are incomplete in that regard.
